---
title: "Cookout Combinatorics"
author: "Joe DiNoto"
date: "4 Oct 2020"
output: 
  html_document: 
    number_sections: yes
    keep_md: yes
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 

 * **For a given food type, what is the most common combination of toppings across two flavor strenghts, limited to people who have two (and only two) flavor strength preferences?**

 * Video link explaining the problem:  https://youtu.be/Wu8Gf963jr8

Scroll down for a peek at the data and an interpretation of the question.

```{r warning=FALSE, message=FALSE}
# packages & loading data
library(tidyverse) 
library(ExPanDaR) # allows for easy permutations
cookout <- read.csv("cookout.csv")  # read csv file
cookout <- cookout %>% tibble() %>%  arrange(Name,Food.Type,Flavor.Strength) 
cookout
```

 
For instance, George prefers bland and medium hamburger toppings. His hamburger flavor combinations acros those two flavor strenghts would be:

 * mayo & onions
 * tomato & onions

Note that mayo & tomato are not a valid combinations because they are both bland flavors. George likes medium and strong ice cream toppings so his ice cream flavor combinations would be

 * cherry & warheads
 * cherry & sour patch
 * oreos & warheads
 * oreos & sour patch

Similarly, cherry & oreos and warheads & sour patch are invalid combinations because they are within the same flavor strength.

# Solution 

## use expand() to find permutations
```{r}
# all unique combinations of names, food types and flavor strengths
# nesting() outputs only combinations that already appear in the data
flavor<- cookout %>% expand(nesting(Name, Food.Type, Flavor.Strength))

# list of names who have 2 flavor strength preferences for hamburgers
hamburger_names <- flavor %>% filter(Food.Type == "hamburger") %>%
  count(Name, sort=TRUE) %>%
  filter(n==2) %>%
  count(Name) %>%
  select(Name)

# list of names who have 2 flavor strength preferences for ice cream
ice_cream_names <- flavor %>% filter(Food.Type == "ice cream") %>%
  count(Name, sort=TRUE) %>%
  filter(n==2) %>%
  count(Name) %>%
  select(Name)

#Filter the list to hamburger names with two flavor strength preferences
hamburger_toppings <- cookout %>% 
  filter(Food.Type == "hamburger") %>%
  filter(Name %in% hamburger_names$Name)

# Filter the list to ice cream names with two flavor strength preferences
ice_cream_toppings <- cookout %>%
  filter(Food.Type == "ice cream") %>%
  filter(Name %in% ice_cream_names$Name)
```

## Flavor strength combinations for Hamburgers

```{r}
hamburger_toppings %>% 
  arrange(Toppings) # sorts toppings a to z so we don't have xy yx problems when counting combinations.

top_vec2<-c() # initializes blank vector

for(k in 1:(nrow(hamburger_toppings)-1)){
  for(i in (k+1):(nrow(hamburger_toppings))){
  if(
    (hamburger_toppings$Name[k]==hamburger_toppings$Name[i]) && 
    # (hamburger_toppings$Food.Type[k] == hamburger_toppings$Food.Type[i]) && 
    (hamburger_toppings$Flavor.Strength[k] != hamburger_toppings$Flavor.Strength[i])
    ){
    z <- data.frame(
      name = c(hamburger_toppings$Name[k]),
      food = c(hamburger_toppings$Food.Type[k]),
      combo = c(paste(hamburger_toppings$Toppings[k],hamburger_toppings$Toppings[i],sep=",")))
    top_vec2 <- rbind(top_vec2,z)
    }
  }
}
```

## Results for hamburgers

```{r}
top_vec2 %>% 
  tibble()%>% 
  group_by(food,combo) %>% 
  summarise(count=n()) %>%
  arrange(desc(count))
```

## Flavor strength combinations for Ice Cream

```{r}
ice_cream_toppings %>% 
  arrange(Toppings) # sorts toppings a to z so we don't have xy yx problems when counting combinations.

top_vec2<-c() # initializes blank vector

for(k in 1:(nrow(ice_cream_toppings)-1)){
  for(i in (k+1):(nrow(ice_cream_toppings))){
  if(
    (ice_cream_toppings$Name[k]==ice_cream_toppings$Name[i]) && 
    # (ice_cream_toppings$Food.Type[k] == ice_cream_toppings$Food.Type[i]) && 
    (ice_cream_toppings$Flavor.Strength[k] != ice_cream_toppings$Flavor.Strength[i])
    ){
    z <- data.frame(
      name = c(ice_cream_toppings$Name[k]),
      food = c(ice_cream_toppings$Food.Type[k]),
      combo = c(paste(ice_cream_toppings$Toppings[k],ice_cream_toppings$Toppings[i],sep=",")))
    top_vec2 <- rbind(top_vec2,z)
    }
  }
}
```

## Results for Ice Cream

```{r}
top_vec2 %>% 
  tibble()%>% 
  group_by(food,combo) %>% 
  summarise(count=n()) %>%
  arrange(desc(count))
```
